Solving Non-smooth Dynamic Problems using \\ the Alternating Direction Method of Multipliers

### 2020-12-07 2021/NODYCON_nonsmooth_ADMM/author.tex
Update NODYCON paper to reflect requests from the reviewers

@@

Another option is to solve the CCP as an optimization problem, using first-order optimization methods such as 
the Nesterov accelerated projected gradient descend  
\cite{hammadTOG2015}
or the Barzilai-Borwein spectral projected gradient
\cite{heynIJNME2013}. These methods are based on a simple projection operator,  matrix-by-vector multiplications and inner products. Their convergence is better than for fixed points iterations, however they share with them the following issue: that they often require the building of a Delassus operator, a matrix that is easy to compute in efficient (factored and sparse) format if diagonal masses are used because these can be easily inverted, but hard to handle if finite elements are added to the problem, because they introduce stiffness and damping matrices whose inverse would be needed as well.

@@

\begin{figure}[h!]
\centering
\includegraphics[width=0.50\textwidth]{t8_convergence.pdf}
\includegraphics[width=0.35\textwidth, trim=0cm -3cm 0 3cm]{t8_snapshot.png}
\caption{Convergence of the ADMMM method within one time step of the wrecking ball benchmark (600 bricks in four walls): residual in frictional constraint violation compared to fixed point Jacobi iterations and to first-order SPG methods.}
\label{fig:t8}
\end{figure}

@@

We performed benchmarks involving multi-body systems with contacts between multiple parts, showing that the performance of the ADMM method is capable of handling problems that would converge too slowly using conventional projected fixed point methods (see Fig.\ref{fig:t8}) or even first-order SPG spectral methods \cite{heynIJNME2013}.

>>>>

Simulation of complex systems involving multiple contacts between mechanical parts can lead to problems of non-smooth dynamics. In such framework, contacts and frictional effects are represented by set-valued functions, thus leading to a more complex mathematical framework respect to conventional smooth Ordinary Differential Equations (ODE) or Differential-Algebraic Equations (DAE). In the latter, in fact, frictional contacts are often approximated as smooth regularizations, hence introducing spring-like penalty forces that allow the adoption of typical ODE or DAE time integrators, but at the same time incurring in low stability unless very small time steps are used.

@@

Another option is to solve the CCP as an optimization problem, using first-order optimization methods such as 
the Nesterov accelerated projected gradient descend  
\cite{hammadTOG2015}
or the Barzilai-Borwein spectral projected gradient
\cite{heynIJNME2013}. These methods are based on a simple projection operator,  matrix-by-vector multiplications and inner products. Their convergence is better than for fixed points iterations, however they share with them the following issue: that they often require the building of a Delassus operator, that is a Schur complement matrix that is easy to compute in efficient (factored and sparse) format if diagonal masses are used because these can be easily inverted, but hard to handle if finite elements are added to the problem, because they introduce stiffness and damping matrices whose inverse would be needed as well.

@@

\begin{figure}[!tbp]
\centering
\includegraphics[width=0.50\textwidth]{t8_convergence.pdf}
\includegraphics[width=0.35\textwidth, trim=0cm -3cm 0 3cm]{t8_snapshot.png}
\caption{Convergence of the ADMMM method within one time step of the wrecking ball benchmark (600 bricks in four walls): residual in frictional constraint violation compared to fixed point Jacobi iterations and to first-order SPG methods.}
\label{fig:t8}
\end{figure}

\subsection{Deformable tire}

\begin{figure}[!tbp]
  \centering
	\includegraphics[width=0.32\textwidth]{tire_5.png}
  \includegraphics[width=0.32\textwidth]{tire_10.png}
	%\includegraphics[width=0.32\textwidth]{tire_15.png}
	\includegraphics[width=0.32\textwidth]{tire_20.png}
	\hfill
  \caption{\textsc{Test 5}. Large deformation of a finite element mesh colliding with a plane. Thanks to the non-smooth formulation, large time steps can be used (only ten steps have been used from start to end) and no penalty parameters must be used in contact points. Contacts between lugs and ground show zero interpenetration regardless of the stiffness of the mesh.}
	\label{fig:tire}
\end{figure}

@@

We performed benchmarks involving multi-body systems with contacts between multiple parts, showing that the performance of the ADMM method is capable of handling problems that would converge too slowly using conventional projected fixed point methods (see Fig.\ref{fig:t8}) or even first-order SPG spectral methods \cite{heynIJNME2013}.
An additional benefit of ADMM, when compared with those methods, is that it does not require the computation of a Schur complement, hence it can be applied seamlessly to problems featuring a block-sparse matrix in place of the diagonal $M$ in \eqref{eq:ChronoMCCP}, such as when introducing stiffness and damping matrices for finite elements (see \ref{fig:tire}).
